### 2020-03-11 Common/chapters/transfer_learning.tex
Julien and Johan: minors

\section{Transfer learning implementation}
\subsection{Layer freezing}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}After each dataset switch, the model is first trained with freezed layers. The freezed layers are those considered as feature extractor, i.e. the first layers of the model. The exact point where the model is split is arbitrary. In our case, it is located after the convolutions. To perform this freezing, multiple steps are required:
\begin{enumerate}
	\item Reset the last layers. To achieve this, an instance of the model is created by loading the state dictionary of the previous one, before replacing the module containing the last layers with a new instance (i.e. a new initialization). 
	\item Iterate over the batch normalization layers and freeze them. Step 3 is not sufficient to freeze batch normalization layers. In an implementation based on PyTorch, this can be achieved by calling \textit{module.eval()} on each module containing such layers (after the \textit{model.train()} call). 
	\item Only pass the parameters of these last layers to the optimizer.
\end{enumerate}

@@

Regarding the full training, the curves contain quite some noise with an already small learning rate of $1*10^{-9}$. An even smaller learning rate of $1*10^{-10}$ was tested out but it turned out to be too small as the training and validation curves were completely flat from the beginning to the end. The noise is also very visible on specific graphs such as the precision and specificity because the Y-axis scale is small. During this last part of the transfer learning, the validation metrics improved during the first 50 epochs approximately, before gently going down as the training metrics kept increasing. At this point, the model reached its optimum. Moreover, the fact that the validation metrics are much higher than the training ones with a dropout probability of 0.0 indicates that the model generalizes well.

\subsection{Visualization of the impact of the various datasets on the target task}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx. As expected, the first training on PROSTATEx generated the same curve as on Figure \ref{fig:tl_ds1_full}. The latter looks steeper because of the difference in the X-axis scale.

Moreover, the frozen training on the second dataset (as well as the other frozen trainings) shows a flat line. Reason for that is that the feature extractor and the decision maker do not change here. In fact, the former is frozen and the latter is always the same. As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training. Furthermore, the full training curves show that brain cancer features are actually helpful. If they were not, the metrics would not have increased as fast and would not have reached such values.

During the full training on the lung dataset, the AUC started at a high value after the first epoch. The combination of the prostate, brain and lung features seem to help the model generalize. This hypothesis is confirmed by the values reached during last training on the prostate dataset. In fact, the AUC value is higher than the one reached during the first training. Interestingly, the curve starts high, improves for 50 epochs and starts decreasing. This implies that the model reached a high level of generalization during transfer learning. These 50 epochs were enough to keep the low-level features learned thanks to the other datasets while adapting the model to the target task again.

@@

\subsection{Conclusion}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning approach using frozen and unfrozen parts allowed to increase the AUC of the model by 15\%. It increased the generalization ability of the model by learning features coming from different body parts. This makes the model more robust and more likely to classify new unseen data well. As this model reached an AUC ($0.80$) similar to the one a less-experienced radiologist can aim for ($0.81$), a few adjustments can create a robust model capable of being used as second diagnosis. These improvements are a step forward in a field where the nonavailability of data is a real obstacle that prevents deep learning-based cancer classification systems from making a major breakthrough.

>>>>

\section{Transfer learning implementation}
\subsection{Layer freezing}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}After each dataset switch, the model is first trained with freezed layers. The freezed layers are those considered as feature extractor, i.e. the first layers of the model. The exact point where the model is split is right after the convolutions. To perform this freezing, multiple steps are required:
\begin{enumerate}
	\item Reset the last layers. To achieve this, an instance of the model is created by loading the state dictionary of the previous one, before replacing the module containing the last layers with a new instance (i.e. a new initialization). 
	\item Iterate over the batch normalization layers and freeze them. Step 3 is not sufficient to freeze batch normalization layers. In an implementation based on PyTorch, this can be achieved by calling \textit{module.eval()} on each module containing such layers (after the \textit{model.train()} call). 
	\item Only pass the parameters of these last layers to the optimizer.
\end{enumerate}

@@

Regarding the full training, the curves contain quite some noise with an already small learning rate of $1*10^{-9}$. An even smaller learning rate of $1*10^{-10}$ was tested out but it turned out to be too small as the training and validation curves were completely flat from the beginning to the end. The noise is also very visible on specific graphs such as the precision and specificity because the y-axis scale is small. During this last part of the transfer learning, the validation metrics improved during the first 50 epochs approximately, before gently going down as the training metrics kept increasing. At this point, the model reached its optimum. Moreover, the fact that the validation metrics are much higher than the training ones with a dropout probability of 0.0 indicates that the model generalizes well.

\subsection{Visualization of the impact of the various datasets on the target task}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx. As expected, the first training on PROSTATEx generated the same curve as on Figure \ref{fig:tl_ds1_full}. The latter looks steeper because of the difference in the x-axis scale.

Moreover, the frozen training on the second dataset (as well as the other frozen trainings) shows a flat line. Reason for that is that the feature extractor and the decision maker do not change here. In fact, the former is frozen and the latter is always the same. As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training. Furthermore, the full training curves show that brain cancer features are actually helpful. If they were not, the metrics would not have increased as fast and would not have reached higher values than the ones reached at DS1 Full.

During the full training on the lung dataset, the AUC started at a high value after the first epoch. The combination of the prostate, brain and lung features seem to help the model generalize. This hypothesis is confirmed by the values reached during last training on the prostate dataset. In fact, the AUC value is higher than the one reached during the first training. Interestingly, the curve starts high, improves for 50 epochs and starts decreasing. This implies that the model reached a high level of generalization during transfer learning. These 50 epochs were enough to keep the low-level features learned thanks to the other datasets while adapting the model to the target task again. Overall, it is noticeable that the maximum AUC is higher at each step.

@@

\subsection{Conclusion}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning approach using frozen and unfrozen parts allowed to increase the AUC of the model by 15\%. It increased the generalization ability of the model by learning features coming from different body parts. This makes the model more robust and more likely to classify new unseen data well. As this model reached an AUC ($0.80$) similar to the one a less-experienced radiologist can aim for ($0.81$), the goal of creating a model which generalizes well is reached. These improvements are a step forward in a field where the nonavailability of data is a real obstacle that prevents deep learning-based cancer classification systems from making a major breakthrough.
